The Castro AMR Simulation Code: Current and Future Developments

### 2019-09-29 astronum_2019_Castro/astronum-castro-2019.tex
updates

\newcommand{\isot}[2]{$^{#2}\mathrm{#1}$}

@@

The recent focus in \castro\ has been on high-fidelity simulations of
reactive flow.  In \cite{castro:sdc} we introduced a new time
integration strategy, spectral deferred corrections (SDC), that
eliminates the coupling error introduced by commonly used operator
splitting techniques (see the discussion in \cite{astronum:2018} for a
graphical illustration of splitting error).  The SDC algorithm used in
\castro\ follows the ideas of \cite{dutt:2000,minion:2003}, and uses
low order explicit advection and implicit reaction updates in a
correction equation that when applied iteratively achieves high-order
time-accuracy.  \castro\ implements both a second-order and
fourth-order method (in space and time) coupling reactions and
hydrodynamics.  At the moment, the method is limited to single levels.

@@

The original approach to parallelism in \castro\ was MPI + OpenMP,
with scaling on manycore architectures achieve using a tiling approach
to OpenMP~\cite{tiling}.

>>>>

\newcommand{\isot}[2]{$^{#2}\mathrm{#1}$}
\newcommand{\avg}[1]{{\left \langle #1 \right \rangle}}
\newcommand{\Rbs}[1]{{\bf R} \left ( #1 \right )}
\newcommand{\Ic}{\bm{\mathcal{I}}}

@@

The recent focus in \castro\ has been on high-fidelity simulations of
reactive flow.  In \cite{castro:sdc} we introduced a new time
integration strategy, spectral deferred corrections (SDC), that
eliminates the coupling error introduced by commonly used operator
splitting techniques (see the discussion in \cite{astronum:2018} for a
graphical illustration of splitting error).  The SDC algorithm used in
\castro\ follows the ideas of \cite{dutt:2000,minion:2003}, and uses
low order explicit advection and implicit reaction updates in a
correction equation that when applied iteratively achieves high-order
time-accuracy.

In an operator split implementation, reaction and burning proceed
without knowing about the other process.  In contrast, the update in
the SDC formulation explicitly couples the processes together.  In \castro\ we consider
reactions and hydrodynamics, and write our conservative system as:
\begin{equation}
\Uc_t = \Advs{\Uc} + \Rbs{\Uc}
\end{equation}
where $\Uc$ is the vector of conserved quantities, $\Advs{\Uc}$ is the
advective term (the divergence of the hydrodynamic fluxes along with
hydrodynamic sources), and $\Rbs{\Uc}$ are the reactive source terms.
The update is done at specific temporal nodes, the number of location
of which are picked to give the desired temporal accuracy.  The update
from one time node $m$ to $m+1$ appears as:
    \begin{align}
      \label{eq:sdc:general}
      \avg{\Uc}^{{m+1},(k+1)} = \avg{\Uc}^{m,(k+1)}
            &+ \delta t_m \left [ \avg{\Advs{\Uc}}^{m,(k+1)} - \avg{\Advs{\Uc}}^{m,(k)} \right ] \nonumber \\
            &+ \delta t_m \left [ \avg{\Rbs{\Uc}}^{{m+1},(k+1)} - \avg{\Rbs{\Uc}}^{{m+1},(k)} \right ] \nonumber \\
            &+ \int_{t^m}^{t^{m+1}} dt  \left (\avg{\Advs{\Uc}}^{(k)} + \avg{\Rbs{\Uc}}^{(k)}\right)
   \end{align}
We use a finite-volume formalism, so $\avg{\Uc}$ is the average of the
conserved state in a zone, $\avg{\Advs{\Uc}}^m$ is the average
advective term in a zone at time $t^m$, and $\avg{\Rbs{\Uc}}^{k+1}$ is
the average of the reactive source at time $t^{m+1}$.  The second
superscript on each term, $(k)$ or $(k+1)$, is the iteration.  This
update is an implicit equation for the new state,
$\avg{\Uc}^{{m+1},(k+1)}$.  The last term in the update is an integral
over the sources constructed from the previous iteration's values at
each time node.  Each iteration of the SDC method increases the
temporal order of accuracy by one, up to the order of accuracy with
which the integral is constructed.  \castro\ implements both a
second-order method (using a trapezoid rule for the integral) and
fourth-order method in space and time (using a Simpsons rule for the
integral and the spatial reconstruction of
\cite{mccorquodalecolella}).  At the moment, the method is limited to
single levels.

@@

The original approach to parallelism in \castro\ was MPI + OpenMP,
with scaling on manycore architectures achieve using a tiling approach
to OpenMP~\cite{tiling}.  More recently we've ported \castro\ to GPUs
using CUDA, using the same computational kernels as the MPI + OpenMP
version.  This reuse keeps the code base managable---we don't need
separate kernels for each architecture---while allowing us to take
advantage of current and next generation supercomputers.  All of the
solvers needed to run our core science problems run on GPUs: the main
unsplit PPM hydrodynamics scheme~\cite{ppm,millercolella:2002}, self-gravity via multigrid with
isolated boundary conditions, thermal diffusion, and nuclear
reactions.  Our approach is to put the data on GPUs and then run all
of the kernels on the GPUs, minimizing data movement.  This has given
us enormous speed-ups.

\begin{figure}[t]
\centering
\includegraphics[width=0.8\linewidth]{wdmerger_gpu}
\caption{\label{fig:gpu} blah}
\end{figure}
